---
title: "OmerOzeren_HMW_2_DATA_607"
output:
  html_document: default
  word_document: default
---

### 3) Copy the introductory example. The vector name stores the extracted names 
```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names
```

**3.1 Use the tools of this chapter to rearrange the vector so that all the elements conform to the standard first_name last_name format.** 
 
In order to get standart first name ,last name we need to remove middle names and titles 

**remove middle names : **
```{r}
names_no_middle_name <- sub(" [A-z]{1}\\. ","",names)
names_no_middle_name
```

**remove titles from the names :**
```{r}
names_final <- sub("[A-z]{2,3}\\. ","",names_no_middle_name)
names_final
```

**DataFrame **
```{r}
df.names <- data.frame(names_final)
df.names
```

**3.2 Construct a logical vector indicating whether a character has a title** 

```{r}
#Recall the original sample 'name2' from part a
titles <- str_detect(names_no_middle_name, "[[:alpha:]]{2,}\\.")
titles
```

```{r}
df.titles<- data.frame(names,titles)
df.titles
```

**3.3 Contruct a logical vector that indicates if a character has a second name**

```{r}
secondname <- str_detect(names, "[A-Z]\\.{1}")
df.secondname <- data.frame(names,secondname)
df.secondname
```
### 4) Describe the types of strings that conform to the following regular expressions and construct an example that is matched by regular expression
**4.1**
 [0-9]+\$

**Any numbers 0-9 zero or more followed by the dollar $  string**
```{r}
sample <- c("5748900000$","omer35$", "38$","38")  
expression = "[0-9]+\\$"
str_detect(sample, expression)
```

**4.2** 
  \\b[a-z{1,4}]\\b

**Any word that has anywhere between 1 to 4 letters** 
```{r}
sample_2 <- c("car","cats","door", "hi", "datascience")
expression_2 <-"\\b[a-z]{1,4}\\b"
str_detect(sample_2, expression_2)
```

**4.3**
  .*?\\.txt$

**Any string that ends with a .txt** 
```{r}
sample_3 <- c("cars.txt", "txt", "timeseries.txt","code3434.txt")
expression_3 <-".*?\\.txt$"
str_detect(sample_3, expression_3)
```

**4.4**
  \\d{2}/\\d{2}/\\d{4}

**Any Numbers that are written in format dd/dd/dddd** 
```{r}
sample_4 <- c("100/1000/10000", "02/12/2019", "2/12/2019")
expression_4 <-"\\d{2}/\\d{2}/\\d{4}"
str_detect(sample_4, expression_4)
```

**4.5**
  <(.+?)>.+?</\\1>

**Text that starts and ends <> with and also  at the end string starts with "/"" ** 
```{r}
sample_5 <- c("<omer>hello</omer>", "<omer>hello<omer>")
expression_5 <-"<(.+?)>.+?</\\1>"
str_detect(sample_5, expression_5)
```

### 9) Extra Credit-The following code hides a secret message. Crack it with R and regular expressions.

```{r}
code <-"clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
code
```

```{r}
#Find all uppercase letters
str_extract_all(code, "[[:upper:]]")