---
title: "Homework 7"
author: "OMER OZEREN"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
    toc_depth: 5
  word_document:
    toc: yes
    highlight: tango
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assignment
Working with HTML, XML and JSON in R

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more than one author. For each book, include the title, authors, and two or three other attributes that you find interesting.

Take the information that you've selected about these three books, and separately create three files which store the book's information in HTML (using an html table), XML, and JSON formats (e.g. "book.html", "book.xml", and "book.json"). To help you better understand the different file structures, I'd prefer that you create each of these files "by hand" unless you're already very comfortable with the file formats.

References to the books.html, books.xml, and books.json.

books.html: https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_7/books.html

books.xml: https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_7/books.xml

books.jason: https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_7/books.json


Load libraries:

```{r message=FALSE, warning=FALSE}
library(RCurl)
library(XML)
library(jsonlite)
library(data.table)
```

### HTML

#### Import HTML file

```{r}
html_url  <- "https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_7/books.html"
html_file <- getURL(html_url)
```

#### Parsing the data
```{r}
html_file <- htmlParse(html_file)
```

#### Converting into a data.frame
```{r}
html_df <- as.data.frame(readHTMLTable(html_file))
html_df
```

### XML

#### Import XML file

```{r}
xml_url <- "https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_7/books.xml"
xml_file <- getURL(xml_url)
```

#### Parsing the data
```{r}
xml_file <- xmlParse(xml_file)
```


#### Converting into a data.frame

```{r}
part_1 <- xmlRoot(xml_file)
xml_df <- xmlToDataFrame(part_1)
xml_df
```

### JSON

#### Import JSON file

```{r}
json_url <- "https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_7/books.json"
json_file <- getURL(json_url)
```

#### Parsing the data
```{r}
json_file <- fromJSON(json_file)
```

#### Converting into a data.frame

```{r}
json_file_df <- as.data.frame(json_file)
json_file_df
```

### Are the three data frames identical?

We can look at the structures of each data.frame to see if they are identical.

```{r}
str(html_df)
```

```{r}
str(xml_df)
```

```{r}
str(json_file_df)
```

After looking at the structures, the HTML and XML data.frame looks identical to each other (other than the names of the headers i.e. Null.Book.Name vs. Book_Name). They are all listed as factors with multiple levels, whereas, the JSON data frame uses chr, int in their data.frame.