---
title: "HomeWork 5"
output: html_document
---

(1) Create a **.CSV** file (or optionally, a **MySQL** database!) that includes all of the information above.  
You're encouraged to use a "wide" structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.

(2) Read the information from your **.CSV** file into **R**, and use **tidyr** and **dplyr** as needed to tidy 
and transform your data.

(3) Perform analysis to compare the arrival delays for the two airlines.

(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative 
descriptions of your data cleanup work, analysis, and conclusions. Please include in your 
homework submission:

The URL to the **.Rmd** file in your GitHub repository and The URL for your **rpubs.com** web page.

# PROCEDURE

## Library definition

```{r library_setups, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}
library(stringr)
library(tidyr)
library(dplyr)
library(knitr)
```


## (1) Read information from **.CSV** file into R.

For simplicity and reproducibility reasons, I have posted this file on my GitHub repository as follows:

### GitHub URL

```{r}
url <- "https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_5/airlines.csv"
```

```{r}
raw_data <- read.csv(url, header=FALSE, sep=",", stringsAsFactors=FALSE)
raw_data <- data.frame(raw_data)
raw_data
```

#### Renaming Column headers

```{r}
# Adding "Missing" titles from original file onto the Row #1
raw_data$V1[1] <- "Airline"
raw_data$V2[1] <- "Status"
# Need to eliminate Row #1 in order to keep data consistency.
#raw_data <-raw_data[-c(1), ]
```

Table  displaying correct column titles.

```{r, echo=FALSE}
kable(raw_data)
```











