---
title: "Homework 5"
author: "OMER OZEREN"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
    toc_depth: 4
  word_document:
    toc: yes
    highlight: tango
    toc_depth: 4
---
---

(1) Create a **.CSV** file (or optionally, a **MySQL** database!) that includes all of the information above.  
You're encouraged to use a "wide" structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.

(2) Read the information from your **.CSV** file into **R**, and use **tidyr** and **dplyr** as needed to tidy 
and transform your data.

(3) Perform analysis to compare the arrival delays for the two airlines.

(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative 
descriptions of your data cleanup work, analysis, and conclusions. Please include in your 
homework submission:

The URL to the **.Rmd** file in your GitHub repository and The URL for your **rpubs.com** web page.



## Library definition

```{r library_setups, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}
library(stringr)
library(tidyr)
library(dplyr)
library(knitr)
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (1) Read information from **.CSV** file into R.

For simplicity and reproducibility reasons, I have posted this file on my GitHub repository as follows:

**GitHub URL**

```{r}
url <- "https://raw.githubusercontent.com/omerozeren/DATA607/master/HMW_5/airlines.csv"
```

```{r}
raw_data <- read.csv(url, header=FALSE, sep=",", stringsAsFactors=FALSE)
raw_data <- data.frame(raw_data)
raw_data
```

## (2) Renaming Column headers

```{r}
# Adding "Missing" titles from original file onto the Row #1
raw_data$V1[1] <- "Airline"
raw_data$V2[1] <- "Status"
# Assigning all the values from the row #1 as the Column Headers
names(raw_data) <- raw_data[1,]
# Need to eliminate Row #1 in order to keep data consistency.
raw_data <-raw_data[-c(1), ]
```

Table  displaying correct column titles.

```{r, echo=FALSE}
kable(raw_data)
```

## (3) Eliminating Empty rows with **"NA"** values 

For this, I have to transform our data as follows:

```{r, echo=FALSE}
# Preview of data composition before transformation
str(raw_data)
```

Procedure to transform values into integers

```{r}
for (i in 3:dim(raw_data)[2]){
  raw_data[,i] <- as.integer(raw_data[,i])
}
```

Preview of data after transformation

```{r, echo=FALSE}
# Preview of data composition before transformation
str(raw_data)
```

Procedure to eliminate all the **NA** lines from our original file by employing **drop_na()**

```{r}
raw_data <- raw_data %>% drop_na()
```

```{r, echo=FALSE}
kable(raw_data)
```

##  (4) Adding missing Airline name to "delayed" row

```{r}
for (i in 1:dim(raw_data)[1]){
  if (i %% 2 == 0){
      raw_data$Airline[i] <-  raw_data$Airline[i-1]
  }
}
```

Final completed table in order to start employing **tidy** transformations for further analysis.

```{r, echo=FALSE}
kable(raw_data)
```

## (5) Analysis


**First:** we need to transform our table by employing **gather()** from **tidyr** library.

```{r}
# Tidy table by having 4 variables (Airline, Status, City, number of flights)
flight <- gather(raw_data, City, Flight_Count, 3:7)
```

```{r, echo=FALSE}
kable((flight))
```

```{r}
# grouping by flights
total_A <- flight %>% group_by(Airline) %>% summarise(Total_Flights = sum(Flight_Count))
kable(total_A)
```

#### (a) Flights Status  by airlines


```{r}
# Total of flights from each airline that were on time
on_time_airline <- flight %>% group_by(Airline) %>% filter(Status == 'on time') %>% summarise(Flights_On_Time = sum(Flight_Count))
kable(on_time_airline)
```

```{r}
# Total of flights from each airline that were delayed.
delayed_airline <- flight %>% group_by(Airline) %>% filter(Status == 'delayed') %>% summarise(Flights_Delayed = sum(Flight_Count))
kable(delayed_airline)
```

**Combine delayed and on_time data sets**
```{r}
# Now will combine all the data set information (including new columns) into data.frame flights.summary
flights_summary_airline <- cbind(on_time_airline, Flights_Delayed = delayed_airline$Flights_Delayed, Total_Flights = total_A$Total_Flights)
flights_summary_airline <- flights_summary_airline %>% mutate(Percent_On_Time_airline = Flights_On_Time/Total_Flights, Percent_Delayed_airline = Flights_Delayed/Total_Flights)

kable(flights_summary_airline)
```

Overall, it appears that AM_West seems to be doing slightly a better job of staying on time. And not to mention, AM West flew more flights than Alaska.

Now I will create two  data.frames where one is Alaska Airlines, and the other is AM_West Airlines.

```{r}
Alaska <- flight %>% filter(Airline == 'ALASKA')
AM_West <- flight %>% filter(Airline == 'AM WEST')
```

**Plot:**
```{r}
my.plot <- barplot(flights_summary_airline$Total_Flights, main="Flights by Airline", xlab="Airline", names.arg=flights_summary_airline$Airline, axes=FALSE, ylim = c(0, max(flights_summary_airline$Total_Flights)+1000))
# Placing values on top of bars
text(my.plot, flights_summary_airline$Total_Flights, labels = flights_summary_airline$Total_Flights, pos = 3)
```

#### (b) Flight Status by City

**grouping by flights**
```{r}
total_C <- flight %>% group_by(City) %>% summarise(Total_Flights = sum(Flight_Count))
kable(total_C)
```

```{r}
# Total of flights from each airline that were on time
on_time_city <- flight %>% group_by(City) %>% filter(Status == 'on time') %>% summarise(Flights_On_Time = sum(Flight_Count))
kable(on_time_city)
```

```{r}
# Total of flights from each airline that were delayed.
delayed_city <- flight %>% group_by(City) %>% filter(Status == 'delayed') %>% summarise(Flights_Delayed = sum(Flight_Count))
kable(delayed_city)
```

**Now will combine all the data set information (including new columns) into data.frame flights.summary**

```{r}
flights_summary_city<- cbind(on_time_city, Flights_Delayed = delayed_city$Flights_Delayed, Total_Flights = total_C$Total_Flights)
flights_summary_city <- flights_summary_city %>% mutate(Percent_On_Time_city = Flights_On_Time/Total_Flights, Percent_Delayed_city = Flights_Delayed/Total_Flights)

kable(flights_summary_city)
```

**Plot:**
```{r}
my.plot <- barplot(flights_summary_city$Total_Flights, main="Flights by Airline", xlab="Airline", names.arg=flights_summary_city$City, axes=FALSE, ylim = c(0, max(flights_summary_city$Total_Flights)+1000))
# Placing values on top of bars
text(my.plot, flights_summary_city$Total_Flights, labels = flights_summary_city$Total_Flights, pos = 3)
```

```{r}
ggplot(flights_summary_city, aes(x = City, y = Percent_On_Time_city)) + geom_point(alpha = 0.5, size = 5, color = 'blue') +  labs(title ="cities on time ratio", x = "City", y = "Flights_On_Time_Percent")
```

```{r}
ggplot(flights_summary_city, aes(x = City, y = Percent_Delayed_city)) + geom_point(alpha = 0.5, size = 5, color = 'blue') +  labs(title ="cities delayed ratio", x = "City", y = "Percent_Delayed")
```

#### (c) Joining tables with horizontal probabilities

```{r}
spread_data <-  flight %>% spread(Status, `Flight_Count`)
```

```{r}
main_table <- spread_data %>% subset(select=c(Airline, City))
airline_table <-  flights_summary_airline %>% subset(select=c(Airline,  Percent_Delayed_airline,Percent_On_Time_airline))
city_table <-  flights_summary_city %>% subset(select=c(City, Percent_Delayed_city,Percent_On_Time_city))
main_table <-  main_table %>% inner_join(airline_table, by="Airline" )
main_table <-  main_table %>% inner_join(city_table, by="City" )
kable(main_table)
```


