---
title: "PROJECT 2"
author: "OMER OZEREN"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
    toc_depth: 5
  word_document:
    toc: yes
    highlight: tango
    toc_depth: 5
---
---
Choose any of the three wide datasets identified in the week 6 discussion items. (You may choose your own)
Read the information from your csv into R and use tidyR and dplyr as needed to transform the data. 
Perform the analysis requested in the discussion item. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries:

```{r results='hide', message=FALSE, warning=FALSE}
library(RCurl)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(psych)
library(knitr)
```

### GINI 

Measuring the wealth distribution between the people in each country has been something economists have been measuring for many years. 
In the GINI index, a higher GINI coefficient signifies inequality in wealth distribution, with 1 being complete inequality and 0 being complete equality.

The World Bank has been maintaining this data.

http://databank.worldbank.org/data/reports.aspx?source=2&series=SI.POV.GINI&country=#

Import dataset from a .csv file.

#### Data
```{r}
GINI.rawfile <- read.csv("https://raw.githubusercontent.com/omerozeren/DATA607/master/PROJECT_2/GINI.csv", header = TRUE)
head(GINI.rawfile)
```


I notice that from the original data frame, there are columns: 1990, 2000, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016,2017,2018. I will use the gather() function from dplyr to 'tidy' up the data.

```{r message=FALSE, warning=FALSE}
#First,  will rename the columns so that it is easier to read.
colnames(GINI.rawfile) <- c("Series.Name", "Series.Code", "Country.Name", "Country.Code", 1990, 2000,  2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016,2017,2018)
# Then  will replace ".." with NA
GINI.rawfile[GINI.rawfile == '..'] <- NA
# Next  will gather the data and then eliminate the columns "Series.Name" and "Series.Code"
GINI <- GINI.rawfile %>% gather(Year, GINI_Index, c(5:16), na.rm = TRUE) %>% group_by(Country.Name) %>% select(-c(Series.Name, Series.Code)) %>% arrange(Country.Name)
#   must convert them into numbers so that we can perform calculations and statistical analysis. 
GINI <- transform(GINI, GINI_Index = as.numeric(GINI_Index))
head(GINI)
```


GINI is now formatted into a 'tidy' format that can now be utilized for analysis.

According to the threat, an analysis that could be performed is the trend in the GINI coefficient for each country (or continent) or the average of the GINI coefficient. 

Let's demonstrate the trend for the countries GINI coefficient scores.

#### Selected countries trend 

```{r warning=FALSE}
GINI.subset <- GINI %>% filter(Country.Name %in% c('United States', 'Vietnam', 'Spain', 'United Kingdom', 'Turkmenistan', 'Russian Federation', 'Portugal', 'Mexico', 'China'))
ggplot(GINI.subset, aes(x = Year, y = GINI_Index)) + geom_jitter(width = 0.5, height = 0.5, aes(color = as.factor(Country.Name))) + geom_line(aes(group = Country.Name), lty = 2, color = "blue") + labs(title = "Subset of Countries", x = "Year", y = "GINI Index") 
```

#### Entire countries trend

```{r warning=FALSE}
ggplot(GINI, aes(x = Year, y = GINI_Index)) + geom_jitter(width = 0.5, height = 0.5, aes(color = as.factor(Country.Name))) + geom_line(aes(group = Country.Name), lty = 2, color = "purple") + labs(title = "Entire Countries", x = "Year", y = "GINI Index") + theme(legend.position = "none")
```

#### Ranked average GINI Index 


```{r}
#  find the max and min of the country's GINI index by dplyr 
par(mfrow = c(1,2))
GINI.avg.per.country.order <- GINI %>% group_by(Country.Name) %>% summarise(AVG_GINI = mean(GINI_Index)) %>% arrange(AVG_GINI)
head(GINI.avg.per.country.order)
GINI.avg.per.country.order.max <- GINI.avg.per.country.order %>% arrange(desc(AVG_GINI))
head(GINI.avg.per.country.order.max)
```

Interestingly, the top 6 counties with the worst GINI indices (higher the number, and hence, worse the inequality) are all in Africa, while the most of the top 6 countries with the best GINI indices are all in Eastern Europe. 

```{r}
GINI.avg.per.country <- GINI %>% group_by(Country.Name) %>% summarise(AVG_GINI = mean(GINI_Index))
GINI.avg.per.country
```

#### Summary 
```{r}
GINI.stat <- describe(GINI.avg.per.country$AVG_GINI)
GINI.qq <- summary(GINI.avg.per.country$AVG_GINI)
GINI.stat
GINI.qq
```

```{r}
hist(GINI.avg.per.country$AVG_GINI, prob = TRUE, main = "GINI Average per Country", xlab = "GINI Index")
x <- seq(20, 70, length = 10000)
y <- dnorm(x, mean = GINI.stat$mean, sd = GINI.stat$sd)
lines(x, y, type = 'l', lwd = 2, col = 'blue')
```

```{r}
qqnorm(GINI.avg.per.country$AVG_GINI)
qqline(GINI.avg.per.country$AVG_GINI)
```

```{r}
# Let's find the GINI mean index value for United States
United_States.GINI <- GINI.avg.per.country[GINI.avg.per.country$Country.Name == 'United States', 'AVG_GINI']
paste0("United_States Average GINI index: ", round(United_States.GINI, 2))
# Calculate the z-score and percentile.
world.mean <- mean(GINI.avg.per.country$AVG_GINI,na.rm = T)
United_States.Z <- (United_States.GINI - world.mean)/GINI.stat$sd
United_States.prob <- pnorm(United_States.Z$AVG_GINI, mean = 0, sd = 1)
paste("United States is", round(United_States.Z$AVG_GINI,2), "standard deviations away from the mean and is", round(United_States.prob, 2) * 100, "percentile.")
```

SUMMARY :
The United Sates has balance out the wealth and equality . The United States is far only  0.3 standard deviation from world mean "GINI SCORE"  and it is  on 62 percentile.

### POPULATION BY COUNTRY

This data shows changes in population by counties  from 1960 to 2017. Here I'm going to define that there have been population shifts and I want to highlight some of them in my analysis in a very easy to read visualization.

The World Bank has been maintaining this data.

https://data.worldbank.org/indicator/sp.pop.totl

Import dataset from a .csv file.

#### Data
```{r}
population<- read.csv("https://raw.githubusercontent.com/omerozeren/DATA607/master/PROJECT_2/populationbycountry.csv", header = TRUE)
head(population)
```

I want to remove Country.Names that actually not country such as 

-North America
-Central & South America
-Antarctica
-Eurasia
-Middle East 
-Asia & Oceania
-World
-Africa
-Europe
-Former Czechoslovakia
-Former Serbia and Montenegro
-Former Yugoslavia
-East
-Hawaiian Trade Zone
-U.S. Pacific Islands
-Wake Island
-Former U.S.S.R.

```{r}
kable(head(population$Country.Name,10))
```

```{r}
#Make a vector that lists all the countries that could not be classified by country code
remove<- c('North America',
           'Central & South America',
           'Antarctica', 
           'Eurasia', 
           'Middle East', 
           'Asia & Oceania', 
           'World', 'Africa', 'Europe', 
           'Former Czechoslovakia',
           'Former Serbia and Montenegro',
           'Former Yugoslavia', 
           'East', 'Hawaiian Trade Zone',
           'U.S. Pacific Islands', 'Wake Island', 'Former U.S.S.R.')
df <- population[ !grepl(paste(remove, collapse="|"), population$Country.Name),]
df <- data.frame(df)
head(df)
```

I need to clean all of my year columns have an X in front of the name. We can use some regular expression to clean the column names. 
```{r}
#remove the x
names(df) <- gsub(x = names(df), pattern = "\\X", replacement = "")  
names(df)
head(df, 2)
```

```{r}
#Lets gather by key value pairs and create a new data frame 
total_pop <- df %>% gather(data=df, Population, "1960":"2017")
head(total_pop, 10)  
tail(total_pop, 10)
```

I need to rename the year column 
```{r}
colnames(total_pop)[colnames(total_pop)=="."]<-"Year"
names(total_pop)
```

removing the Na rows

```{r}
total_pop<-na.omit(total_pop)
head(total_pop)
```

#### subset Population

Creating Subset 
```{r}
df.Turkey<-subset(total_pop, Country.Name=='Turkey', select=c(Country.Name, Year, Population))
head(df.Turkey, 30)
```

```{r}
#Visualize the population
ggplot(data=df.Turkey, aes(x=Year, y=Population, group=1)) +
  geom_line(arrow = arrow())+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs( x="Year", y="Turkey population (In Millions)")
```

#### Entire country Population


```{r}
#Visualize the population
ggplot(data=total_pop, aes(x=Year, y=Population, group=1)) +
  geom_line(arrow = arrow())+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs( x="Year", y="World population (In Millions)")
```

#### Ranked average GINI Index 


```{r}
#  find the max and min of the country's GINI index by dplyr 
par(mfrow = c(1,2))
POP.avg.per.year.order <- total_pop %>% group_by('Year') %>% summarise(AVG_GINI = mean('POPULATION')) %>% arrange(AVG_GINI)
head(GINI.avg.per.country.order)
GINI.avg.per.country.order.max <- GINI.avg.per.country.order %>% arrange(desc(AVG_GINI))
head(GINI.avg.per.country.order.max)
```

Interestingly, the top 6 counties with the worst GINI indices (higher the number, and hence, worse the inequality) are all in Africa, while the most of the top 6 countries with the best GINI indices are all in Eastern Europe. 

```{r}
GINI.avg.per.country <- GINI %>% group_by(Country.Name) %>% summarise(AVG_GINI = mean(GINI_Index))
GINI.avg.per.country
```

